\def\encoding{UTF-8}
\input{mmd-article-header}
\def\mytitle{Homework and Quiz Schedule}
\def\myauthor{Math 352 Differential Equations}
\def\htmlheaderlevel{1}
\def\latexmode{article}
\input{mmd-article-begin-doc}
\chapter{Math 352 Differential Equations---Homework, reading, and quizzes}
\label{math352differentialequationsbr-------homeworkreadingandquizzes}

\section{Daily homework and reading}
\label{daily}

Looking for the weekly problem sets (\autoref{weekly}) or the quizzes (\autoref{quizzes})? Or the \href{sage.html}{Sage tutorial}\footnote{\href{sage.html}{sage.html}}? You can also \href{HW.pdf}{download an experimental PDF}\footnote{\href{HW.pdf}{HW.pdf}} of this homework page.

\subsection{Daily Problems, Week 12: May 6--May 10}
\label{dailyproblemsweek12:may6--may10}

\begin{itemize}
\item Friday, May 10: Multiple spring-mass; vibrating strings; matrix exponentials; loose ends.

\item Wednesday, May 8: $2 \times 2$ systems, phase portraits, saddle points, nodes. Complex eigenvalues.

\item Monday, May 6: Introduction to $ 2 \times 2 $ systems. Download the \href{SystemsOfEquations.pdf}{slides}\footnote{\href{SystemsOfEquations.pdf}{SystemsOfEquations.pdf}}.

\end{itemize}

\subsection{Daily Problems, Week 11: April 29--May 3}
\label{dailyproblemsweek11:april29--may3}

\begin{itemize}
\item Friday, May 3: Advanced matrix theory: eigenvalues, eigenvectors, eigenspaces. Download the \href{MatrixVectorEquations.pdf}{slides}\footnote{\href{MatrixVectorEquations.pdf}{MatrixVectorEquations.pdf}}.

\item Wednesday, May 1: Matrices in Sage: row operations, echelon form, augmented matrices, systems of linear equations. The Fundamental Theorem of Algebra. Download the \href{LinearAlgebraSage.pdf}{lecture notes}\footnote{\href{LinearAlgebraSage.pdf}{LinearAlgebraSage.pdf}} (no slides for today).

\item Monday, April 29: Exam 3. Sections 3.5--3.8.

\end{itemize}

\subsection{Daily Problems, Week 10: April 22--April 26}
\label{dailyproblemsweek10:april22--april26}

\begin{itemize}
\item Friday, April 26: Matrix arithmetic and algebra. Read 7.2; skim 7.1. Download the \href{IntroMatrixAlgebra.sws}{Sage worksheet}\footnote{\href{IntroMatrixAlgebra.sws}{IntroMatrixAlgebra.sws}}.

\begin{itemize}
\item There will probably be homework assigned from outside the textbook.

\end{itemize}

\item Wednesday, April 24: Forced, damped vibrations. Read section 3.8 thoroughly, including all examples and discussion. Introduction to systems of linear differential equations. Skim section 7.2, but don't worry if much seems unfamiliar. We'll augment the text.

\begin{itemize}
\item Section 3.7 (quiz 4\slash 26): 14.

\item Section 3.8 (quiz 4\slash 26): 6, 11.

\end{itemize}

\item Monday, April 22: No class.

\end{itemize}

\subsection{Daily Problems, Week 9: April 15--April 19}
\label{dailyproblemsweek9:april15--april19}

\begin{itemize}
\item Friday, April 19: Forced oscillations, resonance.

\item Wednesday, April 17: Sinusoidal vibrations in standard form: amplitude, frequency, and phase shift. Forced oscillations.

\item Monday, April 15: Free harmonic oscillators. We discussed the undamped and damped cases of the harmonic oscillator problem in one dimension. Slides are forthcoming.

\begin{itemize}
\item Section 3.7 (quiz 4\slash 19): 6, 11 (find $ u $ only), 12 (refer to text)

\end{itemize}

\end{itemize}

\subsection{Daily Problems, Week 8: April 8--April 12}
\label{dailyproblemsweek8:april8--april12}

\begin{itemize}
\item Friday, April 12: Introduced ideas from 3.7 about mechanical vibrations. \href{Quiz8_Solution_M352_S13.pdf}{Solutions}\footnote{\href{Quiz8_Solution_M352_S13.pdf}{Quiz8\_Solution\_M352\_S13.pdf}} to Quiz 8 on undetermined coefficients. Variation of parameters \href{GWVariationParameters_M352_S13.pdf}{group work}\footnote{\href{GWVariationParameters_M352_S13.pdf}{GWVariationParameters\_M352\_S13.pdf}} (\href{GWVariationParameters_Solutions_M352_S13.pdf}{solutions}\footnote{\href{GWVariationParameters_Solutions_M352_S13.pdf}{GWVariationParameters\_Solutions\_M352\_S13.pdf}}) due Wednesday the 17th, not Monday as indicated.

\begin{itemize}
\item Section 3.6 (quiz 4\slash 19): 1--4, 5, 7, 9. Pay special attention to 7.

\end{itemize}

\item Wednesday, April 10: Variation of parameters group work. We practiced some problems from the text.

\item Monday, April 8 (quiz 4\slash 12): Conclusion of Friday's group work (\href{GWUndeterminedCoefficients_Solutions_M352_S13.pdf}{solutions}\footnote{\href{GWUndeterminedCoefficients_Solutions_M352_S13.pdf}{GWUndeterminedCoefficients\_Solutions\_M352\_S13.pdf}}) and wrap-up of undetermined coefficients; introduction to variation of parameters. You can view the presentation in both the usual \href{UndeterminedCoefficientsII.pdf}{PDF format}\footnote{\href{UndeterminedCoefficientsII.pdf}{UndeterminedCoefficientsII.pdf}} and in an \href{UndeterminedCoefficientsII.html}{experimental HTML format}\footnote{\href{UndeterminedCoefficientsII.html}{UndeterminedCoefficientsII.html}} that may be better for mobile devices. Same goes for the presentation on variation of parameters (\href{VariationParameters.pdf}{PDF}\footnote{\href{VariationParameters.pdf}{VariationParameters.pdf}}; \href{VariationParameters.html}{HTML}\footnote{\href{VariationParameters.html}{VariationParameters.html}}).

\end{itemize}

\subsection{Daily Problems, Week 7: April 1--April 5}
\label{dailyproblemsweek7:april1--april5}

\begin{itemize}
\item Friday, April 5: Read section 3.5 and skim 3.6. We had a \href{GWUndeterminedCoefficients_M352_S13.pdf}{group activity}\footnote{\href{GWUndeterminedCoefficients_M352_S13.pdf}{GWUndeterminedCoefficients\_M352\_S13.pdf}} today.

\begin{itemize}
\item Section 3.5: Exercises 1--8.

\end{itemize}

\item Wednesday, April 3: Exam 2. It covers sections 2.5 through 3.4. The best way to review is to work problems (all the way through, not just looking at them). Here are some suitable review problems. As always, they do not constitute any kind of promise about exam content. You are responsible for all the material in the indicated sections, with one exception: the last part of section 2.6, concerning integrating factors for exact equations.

\begin{itemize}
\item 2.5: 9, 13, 15, 22.

\item 2.6: 1–15.

\item 2.7: 1--14; make sure you can execute Euler's method by hand (calculators ok).

\item 3.1: 9–22.

\item 3.2: 1–15.

\item 3.3: 1–22.

\item 3.4: 1–14.

\end{itemize}

\item Monday, April 1: no daily problems. We had a short presentation on undetermined coefficients (the topic of section 3.5). You can view the presentation in both the usual \href{UndeterminedCoefficientsI.pdf}{PDF format}\footnote{\href{UndeterminedCoefficientsI.pdf}{UndeterminedCoefficientsI.pdf}} and in an \href{UndeterminedCoefficientsI.html}{experimental HTML format}\footnote{\href{UndeterminedCoefficientsI.html}{UndeterminedCoefficientsI.html}} that may be better for mobile devices.

\end{itemize}

\subsection{Daily Problems, Week 6: March 18--March 22}
\label{dailyproblemsweek6:march18--march22}

\begin{itemize}
\item Assigned Friday, March 22. \href{ReductionOfOrder.pdf}{Here}\footnote{\href{ReductionOfOrder.pdf}{ReductionOfOrder.pdf}} is today's presentation.

\item Assigned Wednesday, March 20. Read section 3.4 carefully, paying special attention to page 168 and to the last subsection \emph{Reduction of Order}. Once again we had a \href{RepeatedRoots.pdf}{presentation}\footnote{\href{RepeatedRoots.pdf}{RepeatedRoots.pdf}}.

\begin{itemize}
\item Section 3.4: Exercises 1--15, odd.

\end{itemize}

\item Assigned Monday, March 18. Skim section 3.4. Quiz Wednesday, March 20 over 3.2., 3.3. You can also view the (short, mostly recap) \href{OscillatorySolutions.pdf}{presentation}\footnote{\href{OscillatorySolutions.pdf}{OscillatorySolutions.pdf}}.

\begin{itemize}
\item Section 3.3: Exercises 8--22, even.

\end{itemize}

\end{itemize}

\subsection{Daily Problems, Week 5: March 11--March 15}
\label{dailyproblemsweek5:march11--march15}

\begin{itemize}
\item Assigned Friday, March 15. Read section 3.3.

\begin{itemize}
\item Section 3.2: Exercises 24--27.

\item Do all of the Exercises in the \href{GWComplexIntro_M352_S13.pdf}{Introduction to Complex Numbers}\footnote{\href{GWComplexIntro_M352_S13.pdf}{GWComplexIntro\_M352\_S13.pdf}} group assignment. You may find the \href{IntroductionComplexNumbers.pdf}{presentation from today's class}\footnote{\href{IntroductionComplexNumbers.pdf}{IntroductionComplexNumbers.pdf}} helpful as well.

\end{itemize}

\item Assigned Wednesday, March 13. Here is Wednesday's presentation about \href{WronskiansGeneralSolutions_M352_S13.pdf}{Wronskians and general solutions to second-order linear homogeneous equations}\footnote{\href{WronskiansGeneralSolutions_M352_S13.pdf}{WronskiansGeneralSolutions\_M352\_S13.pdf}}.

\begin{itemize}
\item Section 3.2: Exercises 1--6, 13, 14.

\end{itemize}

\item Assigned Monday, March 11. Quiz Friday, March 15 over 2.7 and 3.1. Here is the \href{GWLinearSecondOrder_M352_S13.pdf}{group worksheet}\footnote{\href{GWLinearSecondOrder_M352_S13.pdf}{GWLinearSecondOrder\_M352\_S13.pdf}} we did about exponential solutions to a second-order linear homogeneous differential equation with constant coefficients and the superposition principle (Theorem 3.2.2 in the text.).

\begin{itemize}
\item Section 3.1: Exercises 1--15 odd.

\end{itemize}

\end{itemize}

\subsection{Daily Problems, Week 4: March 4--March 8}
\label{dailyproblemsweek4:march4--march8}

\begin{itemize}
\item Assigned Friday, March 8. You can download the \href{EulersMethod.sws}{Sage worksheet}\footnote{\href{EulersMethod.sws}{EulersMethod.sws}} showing how to implement Euler's Method as a loop. You'll need to upload the \texttt{.sws} file to the Sage server to open it in your account.

\begin{itemize}
\item Section 2.5: Exercises 3--5, 7, 10.

\item Section 2.6: Exercises 1--3, 7.

\item Section 2.7: Exercises 1a, 3a.

\end{itemize}

\item March 6: Section 2.6. No problems today.

\item March 4: Exam 1.

\end{itemize}

\subsection{Daily Problems, Week 3: February 25--March 1}
\label{dailyproblemsweek3:february25--march1}

\begin{itemize}
\item No daily problems for March 1. Exam 1 is coming up! It covers sections 2.1 through 2.4.

\item Assigned Wednesday, February 27. Read section 2.5.

\begin{itemize}
\item Complete the \href{GWModeling_M352_S13.pdf}{activity problems from in-class}\footnote{\href{GWModeling_M352_S13.pdf}{GWModeling\_M352\_S13.pdf}} to be handed in Friday in lieu of Quiz 4. No new daily problems from the text.

\end{itemize}

\item Assigned Monday, February 25. Read section 2.4 and skim 2.5.

\begin{itemize}
\item Section 2.4: Exercises 1--6, 7, 8. No quiz Wednesday.

\end{itemize}

\end{itemize}

\subsection{Daily Problems, Week 2: February 18--22}
\label{dailyproblemsweek2:february18--22}

\begin{itemize}
\item Assigned Friday, February 22. Read section 2.3 and skim section 2.4. Work exercises 2.3.1--2.3.10.

\item Assigned Wednesday, February 20. Exercises from class: 2.2.12, 22, 24, 28. Daily homework is to use Sage or another computer algebra system to plot the solutions to the IVPs. Use the plots to answer the questions about max\slash min and domains of solutions.

\item Assigned Monday, February 18. Read section 2.2 and skim section 2.3. You can \href{SeparationOfVariables.sws}{download}\footnote{\href{SeparationOfVariables.sws}{SeparationOfVariables.sws}} Monday's in-class Sage worksheet; ignore everything after the command consisting only of a zero. You'll need to save the \texttt{.sws} file to disk. Then, you can upload into your own Sage notebook. See the \href{sage.html}{Getting Started worksheet}\footnote{\href{sage.html}{sage.html}} for more.

\begin{itemize}
\item Section 2.2: Exercises 1--8 (all exercises) and 9--15 odd.

\end{itemize}

\end{itemize}

\subsection{Daily Problems, Week 1: February 11--15}
\label{dailyproblemsweek1:february11--15}

\begin{itemize}
\item Assigned Friday, February 15. Read section 2.1 and skim section 2.2.

\begin{itemize}
\item Section 2.1: Exercises 1--5 and 16. To produce the plots, I suggest you make use of \href{sage.html}{Sage}\footnote{\href{sage.html}{sage.html}}. In solving the differential equations themselves, you'll need to make heavy use of integration by parts.

\end{itemize}

\item Assigned Wednesday, February 13. Read section 2.1.

\begin{itemize}
\item Section 1.3: Exercises 7--9, 11, 12, 14. Work done in \href{sage.html}{Sage}\footnote{\href{sage.html}{sage.html}} is fine (except 14), but if you feel you are a little rusty with integrals, you could probably use the practice of doing these by hand. If some of the functions involved look unfamiliar, Google them.

\end{itemize}

\item Assigned Monday, February 11. Read Chapter 1. You're invited to download \href{SlopeFields.sws}{this Sage worksheet}\footnote{\href{SlopeFields.sws}{SlopeFields.sws}} and open it at the public \href{http://www.sagenb.org/}{Sage notebook server}\footnote{\href{http://www.sagenb.org/}{http:/\slash www.sagenb.org\slash }} (registration required) to experiment with slope fields. See: \href{sage.html}{Getting Started with Sage}\footnote{\href{sage.html}{sage.html}}.

\begin{itemize}
\item Section 1.1: Exercises 15--20, 21, 22.

\item Section 1.2: Exercises 5--8, 12--14.

\end{itemize}

\end{itemize}

\section{Weekly homework}
\label{weekly}

Looking for the daily homework (\autoref{daily})?

\subsection{Weekly 6: April 8--April 12}
\label{weekly6:april8--april12}

Due Monday, April 15.

\begin{itemize}
\item Section 3.4: 28.

\item Section 3.5: 9, 10.

\end{itemize}

\subsection{Weekly 5: March 11--March 15}
\label{weekly5:march11--march15}

Due Wednesday, March 20.

\begin{itemize}
\item Section 2.7: 15, 19 (use of Sage is encouraged).

\item Section 3.1: 20, 24.

\end{itemize}

\subsection{Weekly 4: March 4--March 8}
\label{weekly4:march4--march8}

Due Wednesday, March 13.

\begin{itemize}
\item Section 2.5: 15. I encourage you to make use of Sage, although this problem should not be as challenging as some of the 2.3 problems to do by hand. It will help a lot if you think about part (a) in verbal context. What is the meaning of $K$? What is the meaning of $\tau$? How do $\alpha$, $\beta$, etc., correspond to numbers in part (a)? Put some effort into understanding the problem before you begin to solve it.

\item Section 2.6: 12, 16.

\end{itemize}

\subsection{Weekly 3: February 25--March 1}
\label{weekly3:february25--march1}

Due Monday, March 4.

\begin{itemize}
\item Section 2.4: exercises 22, 23, 25, 26.

\end{itemize}

\subsection{Weekly 2: February 18--February 22}
\label{weekly2:february18--february22}

Due Monday, February 25.

\begin{itemize}
\item Section 2.2: exercises 10--16 (even; use Sage or equivalent for plots)

\item Section 2.3: exercises 13, 18, 19.

\end{itemize}

\subsection{Weekly 1: February 11--February 15}
\label{weekly1:february11--february15}

Due Monday, February 18.

\begin{itemize}
\item Section 2.1: exercises 13, 15, 17, 19.

\end{itemize}

\section{Quiz dates}
\label{quizzes}

\begin{itemize}
\item Quiz 10: Friday, April 26 (sections 3.7, 3.8).

\item Quiz 9: Friday, April 19 (sections 3.6, 3.7).

\item Quiz 8: Friday, April 12 (section 3.5).

\item Quiz 7: Wednesday, March 20 (sections 3.2, 3.3).

\item Quiz 6: Friday, March 15 (sections 2.7, 3.1).

\item Quiz 5: Wednesday, March 13 (sections 2.5, 6).

\item Quiz 4: Monday, February 25 (section 2.3).

\item Quiz 3: Wednesday, February 20 (section 2.2). This and all subsequent quizzes will be open-note, but no books or calculators.

\item Quiz 2: Friday, February 15.

\item Quiz 1: Wednesday, February 13.

\end{itemize}

\input{mmd-memoir-footer}

\end{document}
