\documentclass[11pt]{exam}
% \usepackage{pslatex}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{xparse}
%\usepackage{graphicx}
%\DeclareGraphicsExtensions{.jpg, .png}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{enumerate}
\usepackage{siunitx}
\usepackage{tikz}
\usetikzlibrary{arrows}
\NewDocumentCommand\N{}{\mathbf{N}}
\NewDocumentCommand\R{}{\mathbf{R}}
\NewDocumentCommand\Z{}{\mathbf{Z}}
\NewDocumentCommand\Q{}{\mathbf{Q}}
\NewDocumentCommand\C{}{\mathbf{C}}
\NewDocumentCommand\Arg{m}{\mathrm{Arg}\;#1}
\RenewDocumentCommand\Re{m}{\mathrm{Re}\;#1}
\RenewDocumentCommand\Im{m}{\mathrm{Im}\;#1}
\NewDocumentCommand\dwrspace{m}{\vspace*{\stretch{#1}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Fill in these!!

\NewDocumentCommand\termID{}{Spring 2013} % FILL IN TERM ID eg `Spring 2012'
\NewDocumentCommand\courseID{}{Mathematics 352} % FILL IN COURSE ID eg `Mathematics 111'
\NewDocumentCommand\assignmentID{}{Exam 1} % FILL IN ASSIGNMENT ID eg `Quiz 3'
\NewDocumentCommand\dateID{}{March 4, 2013} % FILL IN DATE

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\firstpageheader{}{}{}
\runningheader{\textbf{\termID}}
 {}
 {\textbf{\courseID}}
 %{\emph{Page \thepage~of \numpages}}
\runningheadrule
\setlength{\parskip}{1ex}
\setlength{\parindent}{0pt}
\pagestyle{head}

\begin{document}
\noindent
\textbf{{\large \courseID \\ \assignmentID}}

\noindent
\dateID; 60 minutes  \hfill Name: \underline{\hspace{3in}} 
% \hfill Name: \underline{\hspace{0.5in}Answers\hspace{2in}}

% \printanswers
\addpoints

\noindent
This exam is closed book; you can use a calculator (\emph{not} a cell-phone calculator) but no other electronic aids or printed references. \emph{If the wording or intent of any question is unclear, please ask me to clarify.} I am not trying to confuse you with the problem statements.

You can use your own paper or the provided blank copy paper. Please write your name on each page you hand in. Show all your reasoning and all pertinent calculations. \emph{Give all answers in exact form. Decimal approximations of any accuracy will not receive full credit.}

When you have finished the exam, place this cover sheet on top of it and fold the packet in half \emph{the long way, with your name facing out}.

\dwrspace{1}


\begin{figure}[h]
\centering
    \begin{tikzpicture}
        \path[draw] (0, 0) rectangle (2.8, 4);
        \path[draw] (0, 4) -- (2.0, 4.7) -- (2.0, 4);
    \end{tikzpicture}
\end{figure}
\dwrspace{1}

\begin{center}
    \gradetable
\end{center}

\dwrspace{1}

\begin{center}
    {\Large \emph{Good luck!}}
\end{center}

\newpage

\begin{questions}  

\question Consider the differential equation
\[
    y' - 3y = -9t.
\]
\begin{parts}
    \part[20] Find the general solution to this differential equation, using any method you like.

\dwrspace{1}

    \part[5] Find the particular solution to the differential equation whose graph passes through the point $(2,13)$.
\end{parts}

\dwrspace{1}

\question[25] Using any method you like, find a solution to the initial value problem
\[
    y' = 2xy^2, \quad y(2) = 1.
\]
(You need not express your answer in the form $y = f(x)$.) Are there more solutions than the one you found? Explain.

\dwrspace{1}

\question A tank holds \num{200} \si{\liter} of brine that contains \num{10} \si{\kilo\gram} of salt. Now pure water runs into the tank at a rate of \num{5} \si[per-mode=symbol]{\liter\per\minute} and the well-stirred mixture drains out of the tank at the same rate.

\begin{parts}
    \part[15] Determine the amount of salt in the tank after \num{30} \si{\minute}.

\dwrspace{1}

    \part[10] How long will it take for the amount of salt to drop to \num{1} \si{\kilo\gram}?
\end{parts}

\dwrspace{1}

\question Say as much as you can about the existence and uniqueness of solutions to the following initial value problems. Do not attempt to find any solutions. Answer in complete sentences.

\begin{parts}
    \part[15] ${\displaystyle y' + 2xy = 0, \quad y(0) = 1}$

\dwrspace{1}

    \part[10] ${\displaystyle y' = e^{2y} \cos{2x}, \quad y(\pi/2) = 0}$
\end{parts}

\dwrspace{1}

\end{questions}

\end{document}