\documentclass[11pt]{exam}
% \usepackage{pslatex}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{xparse}
%\usepackage{graphicx}
%\DeclareGraphicsExtensions{.jpg, .png}
\usepackage{amsmath}
%\usepackage{amsfonts}
\usepackage{fourier,MnSymbol}
\usepackage{enumerate}
\usepackage{siunitx}
\usepackage{tikz}
\usetikzlibrary{arrows}
\NewDocumentCommand\N{}{\mathbf{N}}
\NewDocumentCommand\R{}{\mathbf{R}}
\NewDocumentCommand\Z{}{\mathbf{Z}}
\NewDocumentCommand\Q{}{\mathbf{Q}}
\NewDocumentCommand\C{}{\mathbf{C}}
\NewDocumentCommand\Arg{m}{\mathrm{Arg}\;#1}
\RenewDocumentCommand\Re{m}{\mathrm{Re}\;#1}
\RenewDocumentCommand\Im{m}{\mathrm{Im}\;#1}
\NewDocumentCommand\dwrspace{m}{\vspace*{\stretch{#1}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Fill in these!!

\NewDocumentCommand\termID{}{Spring 2013} % FILL IN TERM ID eg `Spring 2012'
\NewDocumentCommand\courseID{}{Mathematics 352} % FILL IN COURSE ID eg `Mathematics 111'
\NewDocumentCommand\assignmentID{}{Exam 2} % FILL IN ASSIGNMENT ID eg `Quiz 3'
\NewDocumentCommand\dateID{}{April 3, 2013} % FILL IN DATE

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\firstpageheader{}{}{}
\runningheader{\textbf{\termID}}
 {}
 {\textbf{\courseID}}
 %{\emph{Page \thepage~of \numpages}}
\runningheadrule
\setlength{\parskip}{1ex}
\setlength{\parindent}{0pt}
\pagestyle{head}

\begin{document}
\noindent
\textbf{{\large \courseID \\ \assignmentID}}

\noindent
\dateID; 60 minutes  \hfill Name: \underline{\hspace{3in}} 
% \hfill Name: \underline{\hspace{0.5in}Answers\hspace{2in}}

% \printanswers
\addpoints

\noindent
This exam is closed book; you can use a calculator (\emph{not} a cell-phone calculator) but no other electronic aids or printed references. \emph{If the wording or intent of any question is unclear, please ask me to clarify.} I am not trying to confuse you with the problem statements.

You can use your own paper or the provided blank copy paper. \textsc{Do not write anything you want graded on the exam paper.} Please write your name on each page you hand in. Show all your reasoning and all pertinent calculations. \emph{Give all answers in exact form. Decimal approximations of any accuracy will not receive full credit.}

When you have finished the exam, place this cover sheet on top of it and fold the packet in half \emph{the long way, with your name facing out}.

\dwrspace{1}

\begin{center}
\textsc{You are strongly encouraged to read all the problems before beginning.}
\end{center}

\dwrspace{1}

% \begin{figure}[h]
% \centering
%     \begin{tikzpicture}
%         \path[draw] (0, 0) rectangle (2.8, 4);
%         \path[draw] (0, 4) -- (2.0, 4.7) -- (2.0, 4);
%     \end{tikzpicture}
% \end{figure}
% \dwrspace{1}

\begin{center}
    \gradetable
\end{center}

\dwrspace{1}

\begin{center}
    {\Large \emph{Good luck!}}
\end{center}

\newpage

\begin{questions}  

\question Find the general solutions to the differential equations.

\begin{parts}
    
    \part[15] $y'' + 3y' - 10y = 0$

    \dwrspace{1}

    \part[15] $y'' + 6y' + 10y = 0$

    \dwrspace{1}

\end{parts}

\question Solve the initial value problems. 

\begin{parts}

    \part[15] $y'' + 3y' - 10y = 0, \quad y(0) = 7, \, y'(0) = -7$

    \dwrspace{1}

    \part[15] $y'' + 6y' + 10y = 0, \quad y(0) = 1, \, y'(0) = -4$

    \dwrspace{1}

\end{parts}    

\question[10] State a condition on the functions $M$ and $N$ that will guarantee that the equation
\[
    M(x, y) + N(x, y) y' = 0
\]
is exact.

\dwrspace{1}

\question[15] Use Euler's method with step size $h = 0.1$ to approximate $y(0.2)$, where $y$ is the unique solution of the initial value problem
\[
    y' = 3y - ty^2, \quad y(0) = 1.
\]
You may express your answer as a decimal, but keep three digits of accuracy if you do. 

\dwrspace{1}

\question[15] Find a fundamental set of solutions of the differential equation $4y'' - 4y' + y = 0$. Justify your answer by computing an appropriate Wronskian.

%\dwrspace{1}
\end{questions}

\end{document}